---
title: "week 4 project"
author: "Jithin Raj R"
date: "August 30, 2017"
output: html_document
---

## peer-graded Assignment: Prediction assignment Write up

```{r}
#reading Test and train data

library(caret)
library(ggplot2)

training <- read.csv(file = "training.csv", na.strings = c("NA","#DIV/0!"))
testing <- read.csv(file = "testing.csv", na.strings = c("NA","#DIV/0!"))


```

### Removing variables with more than 60% NA values

```{r}

#removing variable with more than 60% NA values
na.sum <-colSums(is.na(training))
na.names <- names(na.sum)[na.sum/nrow(training) >.6]
training <- training[,setdiff(colnames(training), na.names)]

#removing variable that are not useful for building model
 training <- training[,-c(1:7)]

```


###  cross validation

- training data set is split into training(75%) and validation set (25%) using createDatapartition()
- 10-fold cross validation is applied using trainControl()

### Expected Out-of-sample error rate 

- Expected out of sample error rate corresponds to accuracy in the test data set. 
- It is the rate of misclassified observation to total observation
- Cross validation data set gives good estimate of out of sample error rate in test data

```{r}

set.seed(1234)
intrain <- createDataPartition(training$classe, p=.75, list = FALSE)

my_train <- training[intrain,]
my_test <- training[-intrain,]
tcontrol <- trainControl(method = "cv", number = 10, preProcOptions = "pca", verboseIter = FALSE, allowParallel = TRUE)


```

## Building Models 

### Random Forest

```{r}
library(rpart)
library(randomForest)

suppressPackageStartupMessages(model1 <- train(y=my_train[,53], x=my_train[,-53] , method="rf",trControl = tcontrol))
rf <- model1

```



```{r}
# validation set is used to generate confusion matrix
pred <- predict(model1, my_test , data=my_test)
confusionMatrix(pred, my_test$classe)
```

####performance of random forest

- Validation set resulted in Accuracy of .9931 and kappa value of 0.9912



### boosted tree

```{r, message= FALSE}

garbage <- capture.output(gbm <- train(y=my_train[,53], x=my_train[,-53] , method="gbm",trControl = tcontrol))
```

```{r}
# validation set is used to generate confusion matrix
gbm.pred <- predict(gbm, my_test)
confusionMatrix(gbm.pred , my_test$classe)

```

#### performance of boosted tree

- Applying Validation data set resulted in accuracy value of 0.9659 and kappa value of .9659


### Predicting using test data

- Random forest model is performing better in than boosted tree. so random forest is used to predict the outcome of      test data

```{r}

pred.test <- predict(rf, testing)
pred.test

```

